---
title: "Geog533 Lab 8"
author: "Amber Jacobson"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}
n <- 24
independent <- 1
regression <- 2.3 - 1.7 
regression
num <- independent
den <- n - 2 
ms1 <- regression/num
res <- 1.7
ms2 <- res/den
fstat <- ms1/ms2

```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|  `r regression` |   `r num`          |  `r ms1`       |     `r fstat`   |
| Residual |       1.7       |  `r den`           |  `r ms2`       |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
a <- .46
b <- .19
x <- 50
y <- a + (b*x)
y
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?

That there is a small correlation between these two elements of data. 

### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
fcrit <- qf(.975, df1 = 1, df2 = 22)

fvalue <- fstat

if(fvalue>fcrit){
  print("We reject the null hypothesis")
} else {
  print("We cannot reject the null hypothesis")
}
```

Therefore the regression coefficient is not significantly different from zero. 


### Question 1(e)
What is the value of the correlation coefficient?
```{r}
total <- 2.3
r2 <- regression/total
r <- sqrt(r2)
r
```





## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
snowfall <- c(36, 78, 11, 45)
elevation <- c(400, 800, 200, 675)
df <- data.frame(snowfall, elevation)
m <- lm(snowfall ~ elevation, data = df)
summary(m)
intercept <- -7.64
slope <- .097 
intercept 
slope
```


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
anova(m)
resmsq <- 126.61
s.error <- sqrt(resmsq)
s.error
```


### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
fcrit2 <- qf(.95, df1=1, df2= 2)
fvalue2 <- 16.173 

if(fvalue2>fcrit2){
  print("We reject the null hypothesis")
} else {
  print("We cannot reject the null hypothesis")
}
```

This means that the regression coefficient associated with independent variables is equal to zero. 

### Question 2(d)
Find the value of $r^2$.
```{r}
result <- cor.test(snowfall, elevation,method = "pearson")
result

r <- result$estimate
r2 <- r^2
round(r2,4)
```


### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}

snowfall <- c(36, 78, 11, 45)
elevation <- c(400, 800, 200, 675)


df <- data.frame(snowfall, elevation)
df$res <- m$residuals
df$predicted <- m$fitted.values
print(df)

```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
n <- 4

regsum <- 2047.77 
ressum <- 253.23 
total <- regsum + ressum 
regdf <- 1 
resdf <- n - 2
msreg <- regsum/regdf 
msred <- ressum/resdf 
fstat2 <- msreg/msred
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|    `r regsum`   |    `r regdf`       |     `r msreg`  |    `r fstat2`   |
| Residual |    `r ressum`   |     `r resdf`      |    `r msred`   |                 |
| Total    |    `r total`    |                    |                |                 |



### Question 2(g)
Graph the data and the regression line.
```{r}
plot(df$elevation, df$snowfall)
abline(m,col="red",lwd=2)
segments(df$elevation,fitted(m),df$elevation,df$snowfall,col="blue",lty = "dashed")
```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
df <- read.csv(url)
m <- lm(df$SalePrice ~ df$LotSize, data = df)
summary(m)
anova(m)

```

The low value of R-squared means there is minimal correlation between the independent and dependent variables. 

## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
url2 = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
df2 <- read.csv(url2)
m2 <- lm(df$SalePrice ~ df$Bedrms, data = df)
summary(m2)
anova(m2)
```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
m3 <- lm(df$SalePrice ~ df$Baths, data = df)
summary(m3)
```

Because R-squared is closer to zero, there is minimal correlation between house price and number of bathrooms. 

## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
hometown <- read.csv(file = "/Users/amberjacobson/Documents/Undergrad/Graduate/533 Statistics/Lab 8/hometown.csv")
hometown 

library(leaflet)

m <- leaflet(width = "100%")
m <- addTiles(m)  
m <- addMarkers(m, lng=42.1642, lat=-75.8335, popup="Amber's Hometown")
m  
# The coordinates are glitching, these are the coordinates that Google Maps provides but it should be Port Crane, NY...not in Antarctica. 



```

